The zoon R package for reproducible and shareable species distribution modelling

### 2017-01-24 ms.Rmd
incorporating Tom's more major comments, including updating Figure 1

The SDM community's inability to compare models and repeat analyses has left the field unable to decisively answer even basic questions about how best to do SDM.
It may also have contributed to widespread misunderstandings about how apply and interpret different modelling approaches [@Yackulic2013].
In order to overcome these problems, the data and methods underpinning SDM research need to be made more accessible, reproducible and modifiable by the whole research community.
This can be achieved if technologies enable and encourage sharing of research as fully reproducible objects [@Peng2011], in ways that suit the diversity of users involved in SDM [@Ahmed2015].

@@

This paper introduces version 0.5.3 of the zoon R package.
We describe the modular structure of zoon workflows and how they can be constructed, shared, reproduced and modified.
We then illustrate how these concepts enable better SDM research by reproducing a published SDM analysis; converting a recently proposed method into a zoon module; and performing a reproducible methodological comparison.

@@

```{r flow_diagram, echo = FALSE, out.width = "375px", fig.cap = "The modular SDM structure encoded by a zoon workflow. A) Description of the five module types. B) Flow diagram illustrating how objects are passed between different module types: `data frame' - an occurrence dataframe; `ras' - a RasterStack object; `mod' - a ZoonModel object; `?' - any user-defined inputs or outputs. C) The flow diagram implied by chaining two `process' modules. D) The flow diagram implied by listing three `model' modules. Full details of module inputs and outputs, and the effects of listing and chaining each module type are given in the zoon vignette `Building a module'."}
knitr::include_graphics("figs/diagrams.png")
```

@@

zoon enables us to investigate this hypothesis by easily comparing MaxEnt models fitted with and without regularisation and feature construction.
We fit MaxEnt with both the usual java implementation as well as the downweighted logistic regression approximation provided in the `maxnet` R package [@maxnet].
We also compare these models with a standard logistic regression model which does not apply the downweighting step.
The following workflow fits these models to a previously published presence-background data on the Caroline wren in the USA [ @maxlike], maps their predictions (Figure 3) and evaluates the predictive performance of each model against the full presence-absence dataset by AUC.
Note that setting the regularisation constant in the `maxnet` package to zero caused numerical errors, so we use a very small value instead.

@@

>>>>

The SDM community's inability to compare models and repeat analyses has left the field unable to decisively answer even basic questions about how best to do SDM.
It may also have contributed to widespread misunderstandings about how apply and interpret different modelling approaches [@Yackulic2013].
In order to overcome these problems, the data and code underpinning SDM research need to be made more accessible, reproducible and modifiable by the whole research community.
This can be achieved if technologies enable and encourage sharing of research as fully reproducible objects [@Peng2011], in ways that suit the diversity of users involved in SDM [@Ahmed2015].

@@

This paper introduces version 0.6 of the zoon R package.
We describe the modular structure of zoon workflows and how they can be constructed, shared, reproduced and modified.
We then illustrate how these concepts enable better SDM research by reproducing a published SDM analysis; converting a recently proposed method into a zoon module; and performing a reproducible methodological comparison.

@@

```{r flow_diagram, echo = FALSE, out.width = "375px", fig.cap = "The modular SDM structure encoded by a zoon workflow. A) Description of the five module types. B) Flow diagram illustrating how objects are passed between different module types: `data frame' - an dataframe of occurrence records; `raster' - a RasterStack object of the covariates; `model' - a ZoonModel object, generating standardised predictions from a given model. C) The flow diagram implied by chaining two `process' modules. D) The flow diagram implied by listing three `model' modules. Full details of module inputs and outputs, and the effects of listing and chaining each module type are given in the zoon vignette `Building a module'."}
knitr::include_graphics("figs/diagrams.png")
```

@@

zoon enables us to investigate this hypothesis by easily comparing MaxEnt models fitted with and without regularisation and feature construction.
We fit MaxEnt with both the widely used java implementation (omitting threshold features) as well as the downweighted logistic regression approximation provided in the `maxnet` R package [@maxnet].
We also compare these models with a standard logistic regression model which does not apply the downweighting step.
The following workflow fits these models to a previously published presence-background data on the Caroline wren in the USA [@maxlike], maps their predictions (Figure 3) and evaluates the predictive performance of each model against the full presence-absence dataset by AUC.
Note that setting the regularisation constant in the `maxnet` package to zero caused numerical errors, so we use a small value instead.

@@

### Supporting Information

Version 0.6 of the zoon R package is archived at https://doi.org/10.5281/zenodo.240926.
The code required to reproduce this manuscript can be downloaded at https://github.com/zoonproject/zoon_app_paper (to be replaced with zenodo archive link after proofing stage).
